---
title: "Excercise 1"
author: "Martin Modrák"
date: "2021/08/24 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: ./libs
    css: ["hygge", "middlebury-fonts", "ninjutsu"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(SBC)
library(ggplot2)

if(packageVersion("SBC") < package_version("0.1.0.0")) {
  stop(paste0("This tutorial has been updated for SBC 0.1, please update your SBC installation"))
}


knitr::opts_chunk$set(echo=FALSE, cache=TRUE, fig.width = 4, fig.height=2.5)
```

# How this tutorial will work

(hopefully)

- Exercise
  - Tasks + questions
  - Bonuses
  - Materials will be available after the tutorial, so you can finish bonuses later.
- Stuck? First ask your colleagues at the table.
  - If you're still stuck, let know any of the organizers.
- It's for you! You can:
  - Mute everything and work alone
  - Take your time
  - Rush through everything

---

# Task 1/1

Introduce yourself to the rest of the table. What projects are you working on?

(3 questions for individual work follow)

---

# Question 1/3

## If SBC signals a problem, does it mean that the model needs fixing?

---

# Question 2/3

## When is a single fit enough to notice a problem?

---

class: split-five

.row[.content[

# Question 3/3
]]
.row[.content[

### Given the rank histogram for a positive variable `sigma` below, which is the rank plot for `log(sigma)`?

]]
.row[.content[

```{r, fig.height=2}
set.seed(3324556)
n_ranks <- 80
max_rank <- 100
ranks_sigma <- data.frame(sim_id = 1:n_ranks, variable = "sigma", rank = round(rbeta(n_ranks, 0.6, 0.9) * max_rank))

plot_rank_hist(ranks_sigma, max_rank = max_rank) + ggtitle(" ")
```


]]
.row[.content[

```{r, fig.height=2}
candidates_log_sigma <- list(
  dplyr::mutate(ranks_sigma, variable = "log(sigma)"),
  data.frame(sim_id = 1:n_ranks, variable = "log(sigma)", rank = round(rbeta(n_ranks, 0.5, 0.9) * max_rank)),
  data.frame(sim_id = 1:n_ranks, variable = "log(sigma)", rank = round(rbeta(n_ranks, 0.9, 0.9) * max_rank))
)

order <- sample.int(3)

plot_rank_hist(candidates_log_sigma[[order[[1]]]], max_rank = max_rank) + ggtitle("A")
plot_rank_hist(candidates_log_sigma[[order[[2]]]], max_rank = max_rank) + ggtitle("B")
plot_rank_hist(candidates_log_sigma[[order[[3]]]], max_rank = max_rank) + ggtitle("C")


```


]]

.row[.content[

&nbsp;



D) can't say, depends on additional properties of `sigma`

]]

---

class: split-three

.row[.content[
# Bonus Question 1

### Match the following rank plots to ECDF plots

Which of those signals a problem? How severe?

]]


.row[.content[

```{r}

set.seed(551324)

max_rank <- 100
N_ranks <- 100

rank_dfs <- list(
  data.frame(variable = "beta", sim_id = 1:N_ranks, rank = round(rbeta(N_ranks, shape1 = 1, shape2 = 1) * max_rank)),
  data.frame(variable = "beta", sim_id = 1:N_ranks, rank = round(rbeta(N_ranks, shape1 = 3, shape2 = 1) * max_rank)),
  data.frame(variable = "beta", sim_id = 1:N_ranks, rank = round(rbeta(N_ranks, shape1 = 8, shape2 = 4) * max_rank))
)

order1 <- sample.int(3)
order2 <- sample.int(3)


plot_rank_hist(rank_dfs[[order1[[1]]]], max_rank = max_rank) + ggtitle("A")
plot_rank_hist(rank_dfs[[order1[[2]]]], max_rank = max_rank) + ggtitle("B") 
plot_rank_hist(rank_dfs[[order1[[3]]]], max_rank = max_rank)  + ggtitle("C")
```


]]
.row[.content[
  
```{r}
plot_ecdf(rank_dfs[[order2[[1]]]], max_rank = max_rank) + ggtitle("1")
plot_ecdf(rank_dfs[[order2[[2]]]], max_rank = max_rank) + ggtitle("2") 
plot_ecdf(rank_dfs[[order2[[3]]]], max_rank = max_rank)  + ggtitle("3")

```
 
  
]]

---

class: split-three

.row[.content[

# Bonus Question 2

### Match the following rank plots to ECDF diff plots

Which of those corresponds to overdispersion/underdispersion? Overestimation/underestimation? 

]]

.row[.content[

```{r}

set.seed(2168855)

max_rank <- 100
N_ranks <- 30

rank_dfs <- list(
  data.frame(variable = "beta", sim_id = 1:N_ranks, rank = round(
    ifelse(runif(N_ranks) < 0.4, rbeta(N_ranks, shape1 = 15, shape2 = 4), rbeta(N_ranks, shape1 = 30, shape2 = 1))
    * max_rank)),
  data.frame(variable = "beta", sim_id = 1:N_ranks, rank = round(rbeta(N_ranks, shape1 = 0.1, shape2 = 0.1) * max_rank)),
  data.frame(variable = "beta", sim_id = 1:N_ranks, rank = round(rbeta(N_ranks, shape1 = 6, shape2 = 6) * max_rank))
)

order1 <- sample.int(3)
order2 <- sample.int(3)


plot_rank_hist(rank_dfs[[order1[[1]]]], max_rank = max_rank) + ggtitle("A")
plot_rank_hist(rank_dfs[[order1[[2]]]], max_rank = max_rank) + ggtitle("B") 
plot_rank_hist(rank_dfs[[order1[[3]]]], max_rank = max_rank)  + ggtitle("C")
```


]]
.row[.content[
  
```{r}
plot_ecdf_diff(rank_dfs[[order2[[1]]]], max_rank = max_rank) + ggtitle("1")
plot_ecdf_diff(rank_dfs[[order2[[2]]]], max_rank = max_rank) + ggtitle("2") 
plot_ecdf_diff(rank_dfs[[order2[[3]]]], max_rank = max_rank)  + ggtitle("3")

```
 
  
]]


