---
title: "SBC Intro"
author: "Martin Modrák"
date: "2021/08/24 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: ./html/libs
    css: ["hygge", "middlebury-fonts", "ninjutsu"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Two types of problems with Stan model

1. Bug in model 

--

2. Data - model mismatch 

--

We will focus on 1.

---

# Simulation to the rescue!

--

How to seperate bugs from model-data mismatch?

--

- If we simulate data _exactly_ as the model assumes, any problem has to be a bug.

---

class: split-two


.column[.content[

## Stan

```stan
data {
   int<lower=0> N;
   vector<lower=0>[N] y;
   vector[N] x;
}

parameters {
   real a;
   real b;
   real<lower=0> sigma;
}

model {
   vector[N] mu = a + b * x;
   a ~ normal(5, 3);
   b ~ normal(0, 1);
   sigma ~ normal(0, 2);
   y ~ lognormal(mu, sigma);
}
```
]]
.column.bg-gray[.content[
  
## R  

```{r, eval = FALSE}
# N and x is not modelled, 
# choose to match your actual data
N <- 10
x <- rnorm(n = N, mean = 0, sd = 2)



a <- rnorm(n = 1, mean = 5, sd = 3)
b <- rnorm(n = 1, mean = 0, sd = 1)
sigma <- abs(
  rnorm(n = 1, mean = 0, sd = 2))


mu <- a + b * x



y <- rlnorm(n = N, 
       meanlog = mu, sdlog = sigma)
       
```  
  
]]

---

# How to spot failure?

```
  variable      mean    sd        q5      q95
   alpha        0.372 0.224   0.00794   0.737 
```

Is this OK, if we simulated

- `alpha = 0.3`

--

-  `alpha = 0.007`

--

-  `alpha = -30`

--

Diagnostics (divergences, Rhat, ESS)

# How to spot succes?!


---
